Particle Interception as a Function of Reynolds Number and Collector Density

### 2020-08-29 code/analysis.Rmd
first version of error prop done
still need version with 20 hz control replaced, and maybe versions with 1 SE error bars instead of 95% CIs

## Objective(s)

@@

### CHANGES FROM VERSION 1 (8/26)

@@

### Time-decay runs

@@

### Velocity experiment

@@

### Flume volume experiment

@@

Here is a function found on StackExchange.com, which performs error propogation in R using derivatives:

@@

Aside from that U-curve, the fit seems reasonably good. Though there are high **absolute** measurements at the beginning, on the **log-scale** they are far less influential than the lower values (`mvc < 10`) at the end. Removing the lower outliers towards the end (`t > 3000`) might be worth trying later if our results don't satisfy.

@@

fitdata <- coefficients(fits) %>%
cbind(date = as.numeric(row.names(coefficients(fits)))) %>%
  left_join(metadata, by = "date")

fitdata %>%
  ggplot(aes(x = dowel_density, y = pump_freq, fill = -t, label = round(-t,7))) +
    geom_tile() + 
    geom_text() +
  scale_fill_gradient(low = "red", high = "white")

@@

### THE PROBLEM IN A NUTSHELL

Having a greater decay rate for the control run than the dowel run, by an amount greater than the difference in $k_s$ between the two runs, implies a negative $k_c$. This seems unrealistic, or at least very out of step with our operating theory. For the rest of this analysis, I'll do what I slipped into my past code and then forgot about: leave out that control run and estimate its values as the mean of the other controls (it does lie at the mean of their independent vars. so weighting the average seems unnecessary). Not ideal! Maybe we should give this set of parameters another shot? But I don't feel great about cherrypicking experimental results that way either.

# THIS CODE WAS USED IN THE FIRST VERSION BUT SHOULD BE LEFT OUT OF ERROR PROPOGATION VERSION

# 
# fitdata[fitdata$date==190802,"t"] <- (fitdata[fitdata$date==190729,"t"] + 
#                                         fitdata[fitdata$date==190321,"t"])/2
# 
# fitdata

@@

hist(trapfinal$sed, breaks = 10, density = 10)

There aren't really any clear outliers, as we established in the boxplots above. While splines might be our final tool, for now an average will do.

final <- trapfinal %>%
  group_by(date) %>%
  summarise(m_trap = mean(sed)/1000) %>% #average sediment in trap converted from mg to g
  mutate(m_s = m_trap*1.95*.6/(pi*.0127^2)) %>% #times test section area divided by area of 1 trap
  left_join(fitdata, by = "date") %>%
  mutate(k_s = m_s/200/(1-exp(t*6000))*(-t))

final %>%
  ggplot(aes(x = dowel_density, y = pump_freq, fill = k_s, label = round(k_s,7))) +
    geom_tile() + 
    geom_text() +
  scale_fill_gradient(low = "red", high = "white")

@@

# THIS CODE WAS USED IN THE FIRST VERSION BUT SHOULD BE LEFT OUT OF ERROR PROPOGATION VERSION

# final[final$date==190802,"k_s"] <- (final[final$date==190729,"k_s"] + 
#                                       final[final$date==190321,"k_s"])/2

Gladly, settling across the dowel treatments does seem to decrease, with a less clear pattern across flow velocity.

@@

bgvals <- final %>%
  filter(dowel_density=="0000") %>%
  transmute(pump_freq, k_bg = -t - k_s) %>%
  arrange(pump_freq)

@@

final <- final %>%
  mutate(k_c = -t - k_s - k_bg)

final %>%
  ggplot(aes(x = dowel_density, y = pump_freq, fill = k_c, label = round(k_c,7))) +
    geom_tile() + 
    geom_text() +
  scale_fill_gradient(low = "red", high = "white")

final %>%
  filter(dowel_density != "0000") %>%
  ggplot(aes(x = as.numeric(pump_freq), y = k_c, color = dowel_density)) +
  geom_line() +
  geom_point()

@@

final <- final %>%
  transmute(frontal_area = as.numeric(dowel_density)*.003175, 
            u = as.numeric(pump_freq)/500, #velocity 
            Re = u*.003175/visc, #Reynolds #
            k_t = -t, 
            k_c, k_s, k_bg) %>%
  mutate(eta = k_c/u/frontal_area) %>%
  mutate(eta = eta * 2.43/(1.95*.4*.6)) # don't forget to correct for time out of test section!

final %>%
  filter(frontal_area != 0) %>%
  ggplot(aes(x = Re, y = eta, color = factor(frontal_area))) +
  geom_line() +
  geom_point()

>>>>

**Objective(s)**

@@

**CHANGES FROM VERSION 1 (8/26)**

@@

**Time-decay runs:**

@@

**Velocity experiment:**

@@

**Flume volume experiment:**

@@

Here is a function found on StackExchange.com, which performs error propogation in R using derivatives. Error values should be named "dx", with x being the corresponding variable name.

@@

Aside from that U-curve, the fit seems reasonably good. **This residual plot was used to remove the most obvious outliers**, though those in the bottom right (say, those satisfying $(y + 0.4) < (0.1/1000)(x-3500)$) might reasonably be considered outliers too. We might be in the gray area in terms of data-massaging at that point. Any formal outlier removal included in the manuscript will have to follow a *de rigueur*, citable method of course.

@@

fitdata <- coefficients(fits) %>%
  mutate(k_t = -t) %>%
  cbind(dk_t = summary(fits)[[4]][,,'t'][,2], date = as.numeric(row.names(coefficients(fits)))) %>%
  left_join(metadata, by = "date")

fitdata %>%
  ggplot(aes(x = pump_freq, y = k_t, ymin = k_t - 1.96*dk_t, ymax = k_t + 1.96*dk_t, color = dowel_density)) +
  geom_line(lty = "dotted", size = 2, alpha = .3) + 
  geom_point(position = position_dodge(width = 3)) +
  geom_errorbar(width = 3, position = "dodge") +
  scale_color_hue(h.start = 0) +
  theme_minimal()

@@

@@

final <- trapfinal %>%
  group_by(date) %>%
  summarise(m_trap = mean(sed)/1000,
            dm_trap = sd(sed)/1000) %>% #average sediment in trap converted from mg to g
  mutate_with_error(m_s ~ m_trap*1.95*.6/(3.14159*.0127^2)) %>% #*A_TS/a_trap
  left_join(fitdata, by = "date") %>%
  mutate_with_error(c_s ~ m_s/200/(1-exp(-k_t*6000))) %>% #intermediate step to squash error 
  mutate_with_error(k_s ~ c_s*k_t)

final %>%
  ggplot(aes(x = pump_freq, y = k_s, ymin = k_s - 1.96*dk_s, ymax = k_s + 1.96*dk_s, color = dowel_density)) +
  geom_line(lty = "dotted", size = 2, alpha = .3) + 
  geom_point(position = position_dodge(width = 3)) +
  geom_errorbar(width = 3, position = "dodge") +
  scale_color_hue(h.start = 0) +
  theme_minimal()

@@

Settling across the dowel treatments does seem to decrease, with a less clear pattern across flow velocity.

@@

bgvals <- final %>%
  filter(dowel_density=="0000") %>%
  mutate_with_error(k_bg ~ k_t - k_s) %>%
  select(pump_freq, k_bg, dk_bg) %>%
  arrange(pump_freq)

@@

final <- final %>%
  mutate_with_error(k_c ~ k_t - k_s - k_bg)

final %>%
  filter(dowel_density != "0000") %>%
  ggplot(aes(x = pump_freq, y = k_c, ymin = k_c - 1.96*dk_c, ymax = k_c + 1.96*dk_c, color = dowel_density)) +
  geom_line(lty = "dotted", size = 2, alpha = .3) + 
  geom_point(position = position_dodge(width = 3)) +
  geom_errorbar(width = 3, position = "dodge") +
  scale_color_hue(h.start = 0) +
  theme_minimal()
```

@@

final <- final %>%
  mutate(frontal_area = as.numeric(dowel_density)*.003175, 
            u = as.numeric(pump_freq)/500, #velocity 
            Re = u*.003175/visc, #Reynolds #
            k_t = -t, 
            k_c, k_s, k_bg) %>%
  mutate(eta = k_c/u/frontal_area,
         deta = dk_c/u/frontal_area) %>%
  mutate(eta = eta * 2.43/(1.95*.4*.6),
         deta = deta * 2.43/(1.95*.4*.6)) # don't forget to correct for time out of test section!

final %>%
  filter(dowel_density != "0000") %>%
  ggplot(aes(x = pump_freq, y = eta, ymin = eta - 1.96*deta, ymax = eta + 1.96*deta, color = dowel_density)) +
  geom_line(lty = "dotted", size = 2, alpha = .3) + 
  geom_point(position = position_dodge(width = 3)) +
  geom_errorbar(width = 3, position = "dodge") +
  scale_color_hue(h.start = 0) +
  theme_minimal()
